---
title: "The MetaRbolomics book"
author: "Jan Stanstrup, Corey D. Broeckling, Rick Helmus, Nils Hoffmann, Ewy Mathé, Thomas Naake, Luca Nicolotti, Kristian Peters, Johannes Rainer, Reza M. Salek, Tobias Schulze, Emma L. Schymanski, Michael A. Stravs, Etienne A Thévenot, Hendrik Treutler, Ralf J. M. Weber, Egon Willighagen, Michael Witting, Steffen Neumann"
always_allow_html: yes
header-includes:
- \usepackage{array}
- \usepackage[table]{xcolor}
- \usepackage{caption}
#urlcolor: blue
bibliography: ["references.bib"]
link-citations: true
---

```{r setup, include=FALSE}
r <- getOption("repos")
r["CRAN"] <- "http://cran.rstudio.com"
options(repos = r)

library(knitr)

format <- opts_knit$get('rmarkdown.pandoc.to')
options(kableExtra.latex.load_packages = FALSE)

if (format=="html") opts_chunk$set(dev = 'svg')
if (format %in% c("pdf", "latex", "epub3"))  opts_chunk$set(dev = 'pdf')
```

```{r table_fun, warning=FALSE, echo=FALSE, message=FALSE}
library(readr)
library(googlesheets)
library(dplyr)

dir.create("public/data", recursive = TRUE)

"https://docs.google.com/spreadsheets/d/1QTe_FhiiFEJJ7hn25V973OV3Y9up8CsiLXgZXgTg_UU/edit?usp=sharing" %>% 
    gs_url() %>% 
    gs_read(1) %>% 
    select(Table, Section, Functionalities, Package, Code_link, Repo) %>% 
    mutate_all(~ifelse(is.na(.),"", .)) %>% 
    write_tsv("public/data/AllMetaRbolomicsTables.csv")

# read package list
pack_list <- read_tsv("public/data/AllMetaRbolomicsTables.csv",
                      col_types=cols(
                                       Table = col_character(),
                                       Section = col_character(),
                                       Functionalities = col_character(),
                                       Package = col_character(),
                                       Code_link = col_character(),
                                       Repo = col_character()
                                     )
                     )


source("scripts/funs.R")
```


`r pagebreak()`

# Preface {-}

The term metaRbolomics has been coined for a workshop at the 2016
annual conference of the international Metabolomics society in Dublin,
Ireland. On May 11, 2016, a small team of authors started to compile a
review of "a few R packages" and how they can be used to do
metabolomics data analysis. After some hiatus, Jan Stanstrup revived
the idea and this turned into a 60+ page review (accepted in MDPI
Metabolites) with more than 200 packages and more than 300
references. But the journey did not end there. In march 2018 Egon
Willighagen suggested that this effort should turn into a book, and
more importantly, should become a community effort, with more
introduction and also include code examples. It now found a home at
[RforMassSpectrometry](https://rformassspectrometry.github.io/RforMassSpectrometry/),
a community website dedicated to R software for the analysis and
interpretation of high throughput mass spectrometry assays, including
proteomics and metabolomics experiments.

## How to ... {-}

**... download data**<br>
The list of packages found in the tables in this book can be downloaded from
[public/data/AllMetaRbolomicsTables.csv](public/data/AllMetaRbolomicsTables.csv).


**... add packages to the tables**<br>
Go to [the googlesheet](https://docs.google.com/spreadsheets/d/1QTe_FhiiFEJJ7hn25V973OV3Y9up8CsiLXgZXgTg_UU/edit?usp=sharing)
and add the package. Please be careful with adding it to the right section.
If it belongs in more than one table add it mulitple times as appropiate.

The package will not appear instantly in the book but only after a change is made to
the book itself. You can also [open an issue](https://github.com/rformassspectrometry/metaRbolomics-book/issues)
and request the reload such that the package shows up.


**... contribute to the text**<br>
There are several options. In order of convenience for the maintainer you can:

* make a pull request on the [GitHub repository](https://github.com/rformassspectrometry/metaRbolomics-book). You will find the text in the `rmd` folder.
* [open an issue](https://github.com/rformassspectrometry/metaRbolomics-book/issues) with the text you want to contribute. Clearly indicate where the text belongs.
* Send your contribution by email to jst( a t )nexs.ku.dk.

Remember to add yourself to author contributions.

## This book in other formats

The book is also available as static [PDF](https://rformassspectrometry.github.io/metaRbolomics-book/The_MetaRbolomics_book.pdf)
file, and as [ePUB](https://rformassspectrometry.github.io/metaRbolomics-book/The_MetaRbolomics_book.epub).


