---
title: "The MetaRbolomics book"
author: "Jan Stanstrup, Corey D. Broeckling, Rick Helmus, Nils Hoffmann, Ewy Mathé, Thomas Naake, Luca Nicolotti, Kristian Peters, Johannes Rainer, Reza M. Salek, Tobias Schulze, Emma L. Schymanski, Michael A. Stravs, Etienne A Thévenot, Hendrik Treutler, Ralf J. M. Weber, Egon Willighagen, Michael Witting, Steffen Neumann"
always_allow_html: yes
header-includes:
- \usepackage{array}
- \usepackage[table]{xcolor}
- \usepackage{caption}
#urlcolor: blue
bibliography: ["references.bib"]
link-citations: true
---

```{r setup, include=FALSE}
r <- getOption("repos")
r["CRAN"] <- "http://cran.rstudio.com"
options(repos = r)

library(knitr)

format <- opts_knit$get('rmarkdown.pandoc.to')
options(kableExtra.latex.load_packages = FALSE)

if (format=="html") opts_chunk$set(dev = 'svg')
if (format %in% c("pdf", "latex", "epub3"))  opts_chunk$set(dev = 'pdf')
```

```{r table_fun, warning=FALSE, echo=FALSE, message=FALSE}
library(readr)
library(googlesheets)
library(dplyr)

dir.create("public/data", recursive = TRUE)

"https://docs.google.com/spreadsheets/d/1QTe_FhiiFEJJ7hn25V973OV3Y9up8CsiLXgZXgTg_UU/edit?usp=sharing" %>% 
    gs_url() %>% 
    gs_read(1) %>% 
    select(Table, Section, Functionalities, Package, Code_link, Repo) %>% 
    mutate_all(~ifelse(is.na(.),"", .)) %>% 
    write_tsv("public/data/AllMetaRbolomicsTables.csv")

# read package list
pack_list <- read_tsv("public/data/AllMetaRbolomicsTables.csv",
                      col_types=cols(
                                       Table = col_character(),
                                       Section = col_character(),
                                       Functionalities = col_character(),
                                       Package = col_character(),
                                       Code_link = col_character(),
                                       Repo = col_character()
                                     )
                     )


source("scripts/funs.R")
```


`r pagebreak()`

# Preface {-}

## How to {-}

**Download data**<br>
The list of packages found in the tables in this book can be downloaded from
[public/data/AllMetaRbolomicsTables.csv](public/data/AllMetaRbolomicsTables.csv).


**Add packages to the tables**<br>
Go to [the googlesheet](https://docs.google.com/spreadsheets/d/1QTe_FhiiFEJJ7hn25V973OV3Y9up8CsiLXgZXgTg_UU/edit?usp=sharing)
and add the package. Please be careful with adding it to the right section.
If it belongs in more than one table add it mulitple times as appropiate.

The package will not appear instantly in the book but only after a change is made to
the book itself. You can also [open an issue](https://github.com/rformassspectrometry/metaRbolomics-book/issues)
and request the reload such that the package shows up.


**Contributing to the text**<br>
There are several options. In order of convenience for the maintainer you can:

* make a pull request on the [GitHub repository](https://github.com/rformassspectrometry/metaRbolomics-book). You will find the text in the `rmd` folder.
* [open an issue](https://github.com/rformassspectrometry/metaRbolomics-book/issues) with the text you want to contribute. Clearly indicate where the text belongs.
* Send your contribution by email to jst( a t )nexs.ku.dk.

Remember to add yourself to author contributions.
