---
author: "Steffen Neumann"
---

`r pagebreak()`

## Appendix 4: miniMetabolomics {-}


```{r setup_mini, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("repos" = list(CRAN="http://cran.rstudio.com/"))

library(miniCRAN)
library(igraph)
library(magrittr)
library(networkD3)

source("scripts/miniMetabolomicsHelper.R")

```

```{r packages}

url="http://www.bioconductor.org/packages/3.9/bioc/"

all_pkgs <- biocAndCranPackages(bioc_url=url)
views <- c("Metabolomics", "Cheminformatics", "Lipidomics", "SystemsBiology")

packagesAndViews <- allPackagesInViews(url)

vpkglist <- lapply(views, function(view) packagesInView(view, packagesAndViews=packagesAndViews))
names(vpkglist) <- views

vpkgs <- unlist(vpkglist)

```

| Repository                 | No. of packages                       |
| -------------------------- | -------------------------------------:|
| CRAN                       | `r table(all_pkgs)["http://cran.rstudio.com/src/contrib"]`   |
| BioC                       | `r table(all_pkgs[,"Repository"])[paste(url, "src/contrib", sep="")]`   |



| View                 | No. of packages                       |
| -------------------- | -------------------------------------:|
| Metabolomics         | `r length(vpkglist[["Metabolomics"]])`   |
| Cheminformatics      | `r length(vpkglist[["Cheminformatics"]])`|
| Lipidomics           | `r length(vpkglist[["Lipidomics"]])`     |
| SystemsBiology       | `r length(vpkglist[["SystemsBiology"]])` |




```{r dependencies}

dg <- makeDepGraph(vpkgs, availPkgs=all_pkgs[vpkgs,], suggests = FALSE, enhances = FALSE)
write.graph(dg, file="metaRbolomics.graphML", format="graphml")

```



```{r denetwork}

# Remove all nodes with fewer than 1 edges
deg <- degree(dg, mode = "in")
idx <- names(which(deg > 0))
gn <- induced.subgraph(dg, idx)

# Extract into data frame and plot
gd <- get.data.frame(gn, what = "edges")
simpleNetwork(gd, fontSize = 12)

```

```{r contract}
gs <- dg

set.seed(42)
# Compute communities (clusters)
cl <- walktrap.community(gs, steps = 5)
cl$degree <- (degree(gs)[cl$names])

# Assign node with highest degree as name for each cluster
cl$cluster <- unname(ave(cl$degree, cl$membership, 
                         FUN=function(x)names(x)[which.max(x)])
)
V(gs)$name <- cl$cluster

# Contract graph ----------------------------------------------------------

# Contract vertices
E(gs)$weight <- 1
V(gs)$weight <- 1
gcon <- contract.vertices(gs, cl$membership, 
                          vertex.attr.comb = list(weight = "sum", name = function(x)x[1], "ignore"))

# Simplify edges
gcon <- simplify(gcon, edge.attr.comb = list(weight = "sum", function(x)length(x)))

## weight threshhold determines how many nodes are shown
gcc <- induced.subgraph(gcon, V(gcon)$weight > 0.5)
V(gcc)$degree <- unname(degree(gcc))

#  ------------------------------------------------------------------------

set.seed(42)
par(mar = rep(0.1, 4)) 
g.layout <- layout.kamada.kawai(gcc)
plot.igraph(gcc, edge.arrow.size = 0.1, layout = g.layout, vertex.size = 0.5 * (V(gcc)$degree))



#  ------------------------------------------------------------------------


# Extract into data frame and plot

library(networkD3)
library(visNetwork)
library(magrittr)

V(gcc)$size <- degree(gcc)

gd <- get.data.frame(gcc, what = "both" )
nodes <- with(gd[["vertices"]],
              data.frame(
                id =  name,
                size =  size
              ))

visNetwork(
  nodes = nodes, 
  edges = gd[["edges"]],
  height = 500,
  width = "100%"
) %>% 
  visPhysics(timestep = 0.03) %>% 
  visOptions(highlightNearest = TRUE)

```

## SessionInfo

This page was created with the following packages:

```{r sessionInfo}
sessionInfo()
```
